---
title: "Project3BellaBeat"
author: "Jade"
date: "2022-09-01"
output: html_document
---

## set working directory (R Script --> Session --> Set Working directory --> select CSV file using read_csv("")) 
Importing Data using readR
Dataset renamed: daily_activity
```{r Importing Data}
library(readr)
daily_activity <- read_csv("dailyActivity_merged.csv")
```

## Initial Summary Information
```{r Summary Information}

colnames(daily_activity)
str(daily_activity)
head(daily_activity)
summary(daily_activity)

```

## Data Cleaning using janitor 

* Column names filtered for characters, numbers, and underscores
* Check data type for ActivityDate column
* Count cells with no values
* Check number of unique Ids
* Check range of dates
* UniqueIdActivityDate has the same number of observations and variables compared to original data set (daily_activity) when accounting for unique Id by ActivityDate
```{r Data Cleaning}

library(janitor)
clean_names(daily_activity)
typeof(daily_activity$ActivityDate)
sum(is.na(daily_activity))

library(tidyverse)
daily_activity %>% group_by(Id)
unique(daily_activity$Id)
daily_activity %>% group_by(ActivityDate)

IdCount <- count(daily_activity, Id)
ActivityDateCount <- count(daily_activity, ActivityDate)
UniqueIdActivityDate <- daily_activity %>% distinct(Id, ActivityDate, .keep_all = TRUE)

```


##  FitBit: Total Steps Per Day
FitBit: Total Steps Per Day
Data collected from 33 individuals from March 12, 2016 to May 12, 2016
Data collected by Amazon survey and submitted to Kaggle by Möbius
```{r}

  ggplot(data = daily_activity, aes(x=ActivityDate, y=TotalSteps, group=1, color = Id)) +
    geom_line()+
    geom_point() +
    facet_wrap(~Id)
    theme(axis.text.x = element_text(angle = 90)) +
    labs(title="FitBit: Total Steps Per Day", subtitle="Data collected from 33 individuals from March 12, 2016 to May 12, 2016", caption="Data collected by Amazon survey and submitted to Kaggle by Möbius")
  
large_color_range <- colorRampPalette(c("blue", "green", "yellow", "orange", "red", "pink", "purple", "black"))  
  new_colors <- large_color_range(33)
  
ggplot(data = daily_activity, aes(x=ActivityDate, y=TotalSteps, group=1, color = Id)) +
    geom_point() +
    scale_colour_gradientn(colors = new_colors) +
    theme(axis.text.x = element_text(angle = 90))

```


## Importing Data using readR
Dataset renamed: sleep

```{r}
library(readr)
sleep <- read_csv("sleepDay_merged.csv")

library(tidyr)
sleepDay <- separate(sleep, 'SleepDay', into=c('Date', 'Time'), sep = ' ')

```


## Creating data sets in environment 
24 Ids
```{r}

entry_perId <- sleepDay %>%
  group_by(Id) %>%
  summarize(sumTotalSleepRecords = sum(TotalSleepRecords))

  # assigned secondary Id (1-24) so the X-axis Ids can be equidistant 
new_entry_perId <- entry_perId %>%
  mutate(new_Id = c(1:24))

  # mean, min, max total number of records per Id
new_entry_perId %>%
  summarize(mean_nrecords = mean(sumTotalSleepRecords), min_nrecords = min(sumTotalSleepRecords), max_nrecords = max(sumTotalSleepRecords))

  # bar graph showing number of sleep records per Id
new_entry_perId %>% 
  ggplot(aes(x = new_Id, y = sumTotalSleepRecords, fill = new_Id)) + 
  geom_col() +
  scale_x_continuous(labels = scales::comma) +
  labs(title="Total Number of Sleep Records Per Id", caption= "mean = 19.2, min = 1, max = 39")


## TotalMinutesAsleep per Date per Id

TotalMinAsleep_perId_perDate <- sleepDay %>%
  group_by(Id, Date) %>%
  summarize(sumTotalMinAsleep = sum(TotalMinutesAsleep))

    # 24 unique Ids
TotalMinAsleep_perId_perDate %>%
  group_by(Id) %>%
  count()

    # ggplot
large_color_range <- colorRampPalette(c("blue", "green", "yellow", "orange", "red", "pink", "purple", "black"))  
new_colors <- large_color_range(24)

## don't really need this one..:   individual geom_point + geom_line + color needed for each Id in the code? to get a proper line graph? Otherwise dataset needs to be organized differently.       
  TotalMinAsleep_perId_perDate %>%
  ggplot(mapping=aes(x= Date, y= sumTotalMinAsleep, color= Id)) +
  geom_point() + 
  scale_colour_gradientn(colors = new_colors) +
  theme(legend.position = "right") +
  theme(axis.text.x = element_text(angle = 80))

## TotalTimeInBed per Date per Id

TotalTimeInBed_perId_perDay <- sleepDay %>%
  group_by(Id, Date) %>%
summarize(sumTotalTimeInBed = sum(TotalTimeInBed))


## don't really need this one..:          
  TotalTimeInBed_perId_perDay %>%
  ggplot(mapping=aes(x= Date, y= sumTotalTimeInBed, color= Id)) +
  geom_point() + 
  scale_colour_gradientn(colors = new_colors) +
  theme(axis.text.x = element_text(angle = 80))


```


## Importing Data using readR
Dataset renamed: weight
8 Ids
```{r}

library(readr)

weight <- read_csv("weightLogInfo_merged.csv")

unique_weightId <- weight %>%
  distinct(weight$Id)

  # number of logs per Id (length is equivalent to COUNT in SQL or COUNTA in Google Sheets)
numlogs_perId <- weight %>%
  group_by(Id) %>%
  summarize(n_of_logs = length(LogId))

  # assigned new Id numbers so X axis is equidistant 
newId_numlogs_perId <- numlogs_perId %>%
  mutate(new_Id = c(1:8))

  # number of logs per Id // geom_col() used when values are contained within data set vs geom_bar which will default count

newId_numlogs_perId %>%
  ggplot(aes(x = new_Id, y = n_of_logs, fill = new_Id)) +
  geom_col() 


```

## Summary comparison among the 3 data sets
```{r}
## 33 daily activity records, 24 sleep records, 8 weight records

variables <- c("daily_activity", "sleep_records", "weight_records")
n_Ids_perlog <- c(33, 24, 8)

group_totals <- data.frame(variables, n_Ids_perlog)

ggplot(data=group_totals) +
  geom_col(mapping=aes(x = variables, y = n_Ids_perlog, fill = variables)) 


```